---
title: "Strings and factors"
output: github_document
---
rvest --data from the web
```{r setup,include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp =0.6,
  out.width = "90%"
) 

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Strings and regex
str_detect --detect pattern --does this pattern appears anywhere in the string--uppercase and lower case matters
```{r}
string_vec = c("my", "name", "is", "jeff")


str_detect(string_vec, "jeff")
str_detect(string_vec, "JEFF")
str_replace(string_vec,"jeff","Jeff")
```


^i think - in the beginning of the line
i think$ - in the end of the line
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```


[Bb]ush --either capital b or lowcase b works
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )
str_detect(string_vec,"bush")
str_detect(string_vec,"Bush")
str_detect(string_vec,"[Bb]ush")
```

number first immediate follow by the letter
^[0-9][a-zA-Z]---anything between 0 and 9, nextone can be a-z or A-Z in the beginning of the line
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

. ---- it is the special character to math any other thing
7.11 -- will also match 7:11, 7-11 (anything between 7 and 11 will match)
7\\.11 -- will only match 7.11
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )
str_detect(string_vec, "7.11")
str_detect(string_vec, "7\\.11")
```

\ is also special character. so we want to use \\.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )
str_detect(string_vec, "\\[")
```

## Factors 
it will have levels in the output
levels : female, male
```{r}
factor_vec = factor(c("male", "male", "female", "female"))
factor_vec
as.numeric(factor_vec)
```

what happens if i relevel...
fct_relevel 
fct_drop --drop the level from the factor
fct_recode -change the label
levels will change to levels : male, female
```{r}
factor_vec = fct_relevel(factor_vec,"male")
factor_vec
as.numeric(factor_vec)
```


## NSDUH--strings
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()

```

- State: pivot anything that is no state 
-str_replace(var,old,new)
-filter i dont want my state variable in  "Total U.S.", "Northeast", "Midwest", "South", "West"
```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


# NSDUH --- factors 
rotate the x variable name -- theme(axis.text.x = element_text(angle = 90, hjust = 1))

```{r}
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
  geom_point() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



